---
title: "Data preparation"
output: html_notebook
---

---

*Fragestellung: * 


*Step Overview:*
1) Drop rows where date 2024+ 
2) Drop all events without id or only with one actor
3) Group actors in general (not top 10)
4) Short to long format (add actor2)
5) 


*Last step : download data to file df_base.csv*



*Libraries*
---

```{r}
# Загрузка необходимых библиотек
library(dplyr)
library(ggplot2)
library(gridExtra)
library(tidyverse)
library(forcats)
library(vcd)  # Установите, если не установлен: install.packages("vcd")
```





*Download data*

```{r}

# Loading the dataset
file_path <- "1997-01-01-2025-01-01-Nigeria.csv"
df_nigeria <- read.csv(file_path, stringsAsFactors = FALSE)

# Displaying the first 5 rows of the dataset
cat("First 5 rows of the dataset:\n")
print(head(df_nigeria, 5))
print(dim(df_nigeria))

```



*1) Drop rows where date 2025+ *

```{r}
df_nigeria <- df_nigeria[df_nigeria$year < 2025, ]
print(dim(df_nigeria))
```


*2) Drop all events without id or only with one actor*

```{r}
# Count occurrences of each event_id_cnty
event_counts <- table(df_nigeria$event_id_cnty)

# Keep only rows where event_id_cnty appears more than once
df_nigeria <- df_nigeria[df_nigeria$event_id_cnty %in% names(event_counts[event_counts > 1]), ]

print(dim(df_nigeria))

```




*3) Group actors in general (not top 10)*


```{r}

df_nigeria$actor1 <- gsub("\\(.*", "", df_nigeria$actor1)
print(head(df_nigeria, 5))

```


*4) Short to long format (add actor2)*

```{r}
library(dplyr)
library(tidyr)  # Load dplyr for easy data manipulation

print("before")
print(dim(df_nigeria))

# Первая таблица: Первое вхождение уникальной комбинации event_id_cnty + actor1
df_first <- df_nigeria %>%
  group_by(event_id_cnty) %>%
  slice(1) %>%
  ungroup()

# Вторая таблица: Второе вхождение уникальной комбинации event_id_cnty + actor1
df_second <- df_nigeria %>%
  group_by(event_id_cnty) %>%
  slice(2) %>%
  ungroup() %>%
  select(event_id_cnty, actor1) %>%  # Берём только нужные колонки
  rename(actor2 = actor1)  # Меняем название, чтобы избежать дублирования

# Объединяем две таблицы по event_id_cnty
df_nigeria <- df_first %>%
  left_join(df_second, by = "event_id_cnty")

# Вывод результата
print(df_nigeria)

print("after")
print(dim(df_nigeria))
```




*Last step*

```{r}
write.csv(df_nigeria, file = "df_base.csv", row.names = FALSE)
df_base <- read.csv("df_base.csv")
```




