---
title: "Data Cleaning"
---


*Steps Overview:*
1) Drop rows where date 2024+ 
2) Unify actors in general (delete year intervals from actors name) 
3) Top 80% + 20% "others" 
  3.1) df_green
4) Add actor2 column
  4.1) df_blue
5) Add actor_pair column
  5.1) df_pink
6) Download df as .csv files




*Libraries*
---

```{r}
# Loading required libraries
# library(dplyr)
library(ggplot2)
library(tidyverse)
library(forcats)
library(vcd)  # Install if not installed: install.packages("vcd")

```





*Download data*

```{r}

# Loading the dataset
file_path <- "1997-01-01-2025-01-01-Nigeria.csv"
df_nigeria <- read.csv(file_path, stringsAsFactors = FALSE)

# Displaying the first 5 rows of the dataset
cat("First 5 rows of the dataset:\n")
print(head(df_nigeria, 5))
print(dim(df_nigeria))

```



*1) Drop rows where date 2025+ *

```{r}
df_nigeria <- df_nigeria[df_nigeria$year < 2025, ]
print(dim(df_nigeria))
```



*2) Unify actors in general (delete year intervals from actors name)*


```{r}

# Remove any text in parentheses (and the parentheses themselves) from the 'actor1' column
df_nigeria$actor1 <- gsub("\\s*\\(.*", "", df_nigeria$actor1)  

# Print the first 5 rows of the df_nigeria dataframe to check the changes
print(head(df_nigeria, 5))  

# Count the number of unique values in the 'actor1' column of df_nigeria
length(unique(df_nigeria$actor1))  

# Count the number of unique values in the 'actor1' column of the df dataframe
length(unique(df$actor1))  

```








*3) Top 85% + 20% "others"*


```{r}
library(dplyr)

# Count the number of mentions for each actor and calculate the percentage
actor_counts <- df_nigeria %>%
  count(actor1, sort = TRUE) %>%
  mutate(percentage = (n / sum(n)) * 100,  # Calculate the percentage of the total
         cumulative_percentage = cumsum(percentage))  # Add cumulative percentage

# Determine which actors fall within 85% of the dataset
top_actors <- actor_counts %>%
  filter(cumulative_percentage <= 85) %>%
  pull(actor1)  # List of actors that make up the top 85%

# Update df_nigeria: replace all other actors with "Others"
df_nigeria <- df_nigeria %>%
  mutate(actor1 = ifelse(actor1 %in% top_actors, actor1, "Others"))

# Print the dimensions of df_nigeria
print(dim(df_nigeria))

```
 
 
 *3.1) df_green*
```{r}
df_green <- df_nigeria
```
 




*5) Add actor2 column*

```{r}
library(dplyr)
library(tidyr)  # Load dplyr for easy data manipulation

print("before")
print(dim(df_nigeria))

# First table: First occurrence of unique combination of event_id_cnty + actor1
df_first <- df_nigeria %>%
  group_by(event_id_cnty) %>%
  slice(1) %>%
  ungroup()

# Second table: Second occurrence of unique combination of event_id_cnty + actor1
df_second <- df_nigeria %>%
  group_by(event_id_cnty) %>%
  slice(2) %>%
  ungroup() %>%
  select(event_id_cnty, actor1) %>%  # Select only the necessary columns
  rename(actor2 = actor1)  # Rename to avoid duplication

# Join the two tables by event_id_cnty
df_nigeria <- df_first %>%
  left_join(df_second, by = "event_id_cnty")

# Replace NA in actor1 with actor2 if actor1 == NA
# Replace NA in actor2 with actor1 if actor2 == NA
df_nigeria <- df_nigeria %>%
  mutate(actor1 = ifelse(is.na(actor1), actor2, actor1),
         actor2 = ifelse(is.na(actor2), actor1, actor2))

# Print the result
print(df_nigeria)

print("after")
print(dim(df_nigeria))

```


*4.1) df_blue*

```{r}
df_blue <- df_nigeria
```




*5) Add actor_pair column*
    *5.1) df_pink*



```{r}
# Sort actor pairs so that (A, B) and (B, A) are considered the same
df_pink <- df_blue %>%
  mutate(actor_min = pmin(actor1, actor2, na.rm = TRUE),  # Get the minimum actor (considering NA)
         actor_max = pmax(actor1, actor2, na.rm = TRUE))  # Get the maximum actor

# Handle cases where (NA, A) → just take A
df_pink <- df_pink %>%
  mutate(actor_pair = ifelse(is.na(actor_min), actor_max,  # If actor_min is NA, take actor_max
                             ifelse(is.na(actor_max), actor_min,  # If actor_max is NA, take actor_min
                                    paste(actor_min, actor_max, sep = " + "))))  # Otherwise, combine both
```







*6) Download df as .csv files*

```{r}
write.csv(df_blue, file = "df_blue.csv", row.names = FALSE)
write.csv(df_green, file = "df_green.csv", row.names = FALSE)
write.csv(df_pink, file = "df_pink.csv", row.names = FALSE)

```



